---
title: "HW_9 CO2"
author: "Amy Hessl"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

You will need to install and load these packages to get this .Rmd to render figures:
```{r, libraries}
library(markdown)
library(knitr)
library(kableExtra)
library(data.table)
```

For this assignment, please use the file _data/co2_emissions_tonnes_per_person_gapminder.csv_
You will collaborate with a partner to complete the homework. As with the previous assignment, each student should complete the assignment first, then you should merge your answers on one student's cloned HW_9 repository.  Finally, the host student should make a pull request to submit your answers. 
Turn in a _.Rmd_ file that is well-formatted and documented.

*Some Hints on Tables in Rmd:*
For fancy styling of tables for html output from .Rmd, try the kable() function in knitr package:
```
kable(my_tab)
```

Or even fancier, use the kablextra package:
```
my_tab %>%
  kable() %>%
  kable_styling()
```


***
### Questions:

1. Are the data in a flat file format? 
Yes
```{r}
co2_emissions_tonnes_per_person <- read.csv(file = "data/co2_emissions_tonnes_per_person_gapminder.csv", header=TRUE, sep = ",") #read in .csv file
```
2. What is the structure of the data?  Are there any variables you might need to convert prior to manipulating the data?
The data is in a data frame. All years are numeric vectors. Names of countries are factors. Might want to change the country vector to character data type. 

```{r, echo=FALSE}
co2_emissions_tonnes_per_person$country <- as.character(co2_emissions_tonnes_per_person$country) #convert country vector to character
```

3. What countries had per capita emissions larger than the 95th percentile in 2014? Make a table of their names and per capita emissions. Hint: Use quantile() to find the 95% cutoff.

```{r,echo=FALSE}
dat <- co2_emissions_tonnes_per_person$X2014>quantile(co2_emissions_tonnes_per_person$X2014,0.95) #countries that are in the top 5% of CO2 producers per capita in 2014
table <- data.frame("Country"=co2_emissions_tonnes_per_person$country[dat], "PerCapEmissions.2014"=co2_emissions_tonnes_per_person$X2014[dat]) #table containing top 5% of CO2 producers per capita and amount of CO2 produced in 2014 (units in tons)
table %>%
  kable() %>%
  kable_styling()
```

4. Make a table comparing the per capita emissions of the US and the world's highest per capita emitter (from above)  in 1925, 1950, and 2014.  
```{r, echo=FALSE}
max <- which.max(co2_emissions_tonnes_per_person$X2014) #gives the factor of the highest emitting country in 2014
TopCountry <- as.character(co2_emissions_tonnes_per_person$country[max]) #converts factor to character
TopCountry
```



```{r, echo=FALSE}
countries <- co2_emissions_tonnes_per_person[co2_emissions_tonnes_per_person$country %in% c("United States", "Qatar"), ]#index United States and Qatar
fdata <- countries[ , c("X2014", "X1950", "X1925")]
ex_tab <- data.frame(fdata)
colnames(ex_tab) <- c("2014", "1950", "1925")
row.names(ex_tab) <- c("Qatar", "United States")
ex_tab %>%
  kable() %>%
  kable_styling()
```

